---
title: "Sensitive question RCT"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

This project aims to assess the impact of asking sensitive questions on application completion rates using GCF data. We test these hypotheses using two list experiments targeting GCF applicants experiencing unstable housing situations or having a non-citizen family member. The sampling frame is limited to those who consent to receive emails or SMS texts.

# import data 

Only the last 2 years since the contact information in the older data would have been redacted by our privacy policy and DSA with the CDSS.

We use the data collected from 2023-10-01 through 2024-04-29.

```{sql}
SELECT *
FROM gcf_public.all_applications
WHERE started_at >= '2023-10-01'
AND 
started_at <= '2024-04-29'
```

```{r}
write_csv(df1, "df1.csv")
```

```{r}
library(tidyverse)
library(janitor)
```

```{r}
df1 <- read_csv("df1.csv")
```

```{r}
add_c4a_theme <- function(ggplot_obj, target = "fill", legend_name = "Legend") {

  message("The current function only has 6 color choices.")

  SN_LIGHT_GREEN <- "#00A88F"
  SN_DEEP_GREEN <- "#006152"
  SN_BLACK <- "#000000"
  SN_DARK_BLUE <- "#2B1A78"

  CFA_DARK_BLUE <- SN_DARK_BLUE
  CFA_DEEP_GREEN <- SN_DEEP_GREEN
  CFA_BLACK <- SN_BLACK

  CFA_RICH_RED <- "#EA3C48"
  CFA_LIGHT_BLUE <- "#A1B4EA"
  CFA_WARM_YELLOW <- "#FFB446"
  CFA_SAND <- "#E9CCBE"

  color_theme <- c(CFA_LIGHT_BLUE, CFA_DARK_BLUE, SN_LIGHT_GREEN, SN_DEEP_GREEN, CFA_WARM_YELLOW, CFA_RICH_RED)

  base <- ggplot_obj +
    theme(
      text = element_text(
        size = 12,
        family = "Helvetica Neue"
      ),
      axis.title.y = element_text(face = "bold", size = 12),
      axis.title.x = element_text(face = "bold", size = 12),
      axis.text.x = element_text(size = 10),
      plot.title = element_text(face = "bold"),
      legend.title = element_text(size = 11)
    )

  if (target == "fill") {
    out <- base +
      scale_fill_manual(name = legend_name, values = color_theme) # the legned is a default name. please feel free to change it.
   }

  if (target == "color") {
    out <- base +
      scale_color_manual(name = legend_name, values = color_theme) # the legned is a default name. please feel free to change it.
  }

  message(paste("cfa theme is applied to", target, "parameter.")) 

  return(out)
}
```

```{r}
df1 <- df1 %>%
    mutate(unstable_housing = ifelse(stable_housing == FALSE, TRUE, FALSE)) 
```

```{r}
df1 %>%
    filter(unstable_housing == TRUE | household_includes_noncitizen == TRUE) %>%
    group_by(unstable_housing, household_includes_noncitizen) %>%
    count() %>%
    ungroup() %>%
    mutate(freq = round(n/sum(n),2))
```

```{r}
df1$opened_date <- as.Date(df1$started_at, format = "%m/%d/%Y")
```

```{r}
df1 %>%
    group_by(opened_date) %>%
    count() %>%
    ggplot(aes(x = opened_date, y = n)) +
    geom_point() +
    ylim(0, 10000) +
    labs(x = "Application start date", 
         y = "Count", 
         caption = "GCF (2023.10.01-2024.4.29)")
```

```{r}
df1 %>%
    mutate(miss_housing_q = ifelse(is.na(stable_housing), "Didn't respond homeless Q", "Responded homeless Q")) %>%
    group_by(miss_housing_q) %>%
    summarize(
        n = n(), 
        dropoff_rate = mean(completed_application == FALSE, na.rm = TRUE),
        lower_ci = mean(completed_application == FALSE, na.rm = TRUE) - qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n()),
        upper_ci = mean(completed_application == FALSE, na.rm = TRUE) + qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n()),
        .groups = 'drop'
    )
```

```{r}
df1 %>%
    mutate(miss_imm_q = ifelse(is.na(household_includes_noncitizen), "Didn't respond non-citizen Q", "Responded non-citizen Q")) %>%
    group_by(miss_imm_q) %>%
    summarize(
        n = n (),
        dropoff_rate = mean(completed_application == FALSE, na.rm = TRUE),
        lower_ci = mean(completed_application == FALSE, na.rm = TRUE) - qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n()),
        upper_ci = mean(completed_application == FALSE, na.rm = TRUE) + qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n()),
        .groups = 'drop'
    )
```

```{r}
df1 %>%
    group_by(household_includes_noncitizen) %>%
    summarize(
        dropoff_rate = mean(completed_application == FALSE, na.rm = TRUE),
        lower_ci = mean(completed_application == FALSE, na.rm = TRUE) - qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n()),
        upper_ci = mean(completed_application == FALSE, na.rm = TRUE) + qt(0.975, df = n() - 1) * sd(completed_application == FALSE, na.rm = TRUE) / sqrt(n())
    )
```

{# housing #}
    , case
        when a.living_situation_couchsurfing
            or a.living_situation_lives_in_vehicle
            or a.living_situation_stays_in_shelter
            or a.living_situation_on_the_streets
            or a.living_situation_other
        then false
        when a.living_situation_rent_or_own
            or a.living_situation_lives_on_campus
        then true
    end as stable_housing
    
```{r}
df1_filtered <- df1 %>%
    filter(stable_housing == FALSE | household_includes_noncitizen == TRUE)

homeless_sample <- df1_filtered %>%
    filter(stable_housing == FALSE & household_includes_noncitizen == FALSE) # only homeless

noncitizen_sample <- df1_filtered %>%
    filter(stable_housing == TRUE & household_includes_noncitizen == TRUE) # only noncitizen
```

# Sampling frame

```{r}
email_contact <- read_csv("members__filtered_by_created_at_and_email_address__sorted_by_created_at_descending_2024-04-29T18_18_06.310356Z.csv") %>%
    clean_names() %>%
    select(calfresh_application_id, email_address)

phone_contact <- read_csv("members__filtered_by_phone_number__created_at__and_phone_number_type__sorted_by_created_at_descending_2024-04-29T18_18_32.982004Z.csv") %>%
    clean_names() %>%
    select(calfresh_application_id, phone_number)

homeless_sample$calfresh_application_id <- as.character(homeless_sample$calfresh_application_id)

email_contact$calfresh_application_id <- as.character(email_contact$calfresh_application_id)

phone_contact$calfresh_application_id <- as.character(phone_contact$calfresh_application_id)

noncitizen_sample$calfresh_application_id <- as.character(noncitizen_sample$calfresh_application_id)
```

```{r}
homeless_sample <- homeless_sample %>%
    left_join(email_contact) %>%
    left_join(phone_contact) %>%
    filter(!(is.na(email_address) & is.na(phone_number))) %>%
    select(calfresh_application_id, email_address, phone_number, county_name)

noncitizen_sample <- noncitizen_sample %>%
    left_join(email_contact) %>%
    left_join(phone_contact) %>%
    filter(!(is.na(email_address) & is.na(phone_number))) %>%
    select(calfresh_application_id, email_address, phone_number, submitted_language)
```

# Power analysis 

Reference for the effect sizes (from table 4 in the following article): https://academic.oup.com/poq/article/77/S1/159/1878470 

```{r}
install.packages("pwr")
library(pwr)

# Install the pwr package if not already installed
if (!requireNamespace("pwr", quietly = TRUE)) {
  install.packages("pwr")
}

# Load the pwr package
library(pwr)

# Calculate the sample size
# effect size = (0.03 + 0.1)/2, significance level alpha = 0.05
sample_size <- pwr.t.test(d = (0.03 + 0.1)/2, 
    sig.level = 0.05, 
    power = 0.8, 
    type = 'two.sample', 
    alternative = 'two.sided')$n

# Assume 10% response rate 
target_size <- sample_size / 0.10

target_size
```

```{r}
paste0(target_size / round(nrow(homeless_sample), 2) * 100, "%", " of the GCF applicants experienced unstable housing");
paste0(target_size / round(nrow(noncitizen_sample), 2) * 100, "%", " of the GCF applicants whose households include a noncitizen member")
```

"28.0726482791893% of the GCF applicants experienced unstable housing"
"102.476080693776% of the GCF applicants whose households include a noncitizen member"


# Sample selection

```{r}
# set seed for computational reproducibility
set.seed(1234)

homeless_sample %>%
    group_by(county_name) %>%
    count() %>%
    ungroup() %>%
    mutate(freq = n/sum(n)) %>%
    arrange(desc(freq)) %>%
    head(10)
```

We grouped the data to create a 30% random sample because there's a chance that applicants from top counties (e.g., Los Angeles, San Diego, and Riverside) will dominate the data.

```{r}
homeless_sample_filtered <- homeless_sample %>%
  group_by(county_name) %>%
  sample_frac(0.3)
```

46% of the GCF applicants submitted their applications written in Spanish.

```{r}
noncitizen_sample %>%
    group_by(submitted_language) %>%
    count() %>%
    ungroup() %>%
    mutate(freq = n/sum(n))
```

# Random assignment

```{r}
homeless_sample_filtered <- homeless_sample_filtered %>%
    mutate(contact_method = case_when(
        is.na(phone_number) & !is.na(email_address) ~ "email",
        !is.na(phone_number) & is.na(email_address) ~ "phone",
        !is.na(phone_number) & !is.na(email_address) ~ "both",
        is.na(phone_number) & is.na(email_address) ~ "no_contact"
    ))

noncitizen_sample <- noncitizen_sample %>%
    mutate(contact_method = case_when(
        is.na(phone_number) & !is.na(email_address) ~ "email",
        !is.na(phone_number) & is.na(email_address) ~ "phone",
        !is.na(phone_number) & !is.na(email_address) ~ "both",
        is.na(phone_number) & is.na(email_address) ~ "no_contact"
    ))

homeless_sample_filtered %>%
    group_by(contact_method) %>%
    count() %>%
    ungroup() %>%
    mutate(freq = n/sum(n))

noncitizen_sample %>%
    group_by(contact_method) %>%
    count() %>%
    ungroup() %>%
    mutate(freq = n/sum(n))
```

```{r}
if (!require(randomizr)) install.packages("randomizr") # for randomization 
library(randomizr)
```

## Homeless blocked by contact method

```{r}
block_ra_homeless <- block_ra(
    blocks = c(homeless_sample_filtered$contact_method), 
    prob_each = c(0.5, 0.5), 
    conditions = c("control", "treatment")
)

homeless_sample_filtered$conditions <- block_ra_homeless
```

## Noncitizen blocked by contact method

```{r}
block_ra_noncitizen <- block_ra(
    blocks = c(noncitizen_sample$contact_method), 
    prob_each = c(0.5, 0.5), 
    conditions = c("control", "treatment")
)

noncitizen_sample$conditions <- block_ra_noncitizen
```

# Export files

```{r}
write_csv(homeless_sample_filtered, "homeless_sample.csv");
write_csv(noncitizen_sample, "noncitizen_sample.csv")

sample <- bind_rows(homeless_sample_filtered %>%
    ungroup() %>%
    select(-c("county_name")) %>%
    mutate(group = "homeless"),
    noncitizen_sample %>%
    select(-c("submitted_language")) %>%
    mutate(group = "immigrant")
)

sample <- sample %>%
    rename(assignment = conditions)
```

```{r}
write_csv(sample, "sample.csv")
```

# Balance checks

```{r}
install.packages("modelsummary")
library(modelsummary)
```

```{r}
sample_covars <- sample %>% 
    select(c("calfresh_application_id", "contact_method", "assignment", "group")) %>%
    left_join(
        df1 %>%
        select(c("calfresh_application_id", "age", "sex", "ssi", "county_name", "household_size", "eligible_household_size", "has_job", "self_employed", "student", "final_eligible_gross_income_amount", "final_eligible_net_income_amount", "submitted_language")),
    by = c("calfresh_application_id"))

datasummary_balance(~assignment, 
    sample_covars %>%
    select(-c("calfresh_application_id", "contact_method")) %>%
    mutate(submitted_language = ifelse(submitted_language == "vi", NA, submitted_language))
    ,
    output = "DT")

datasummary_balance(~assignment, 
    sample_covars %>%
    select(-c("calfresh_application_id", "contact_method"))
    ,
    output = "latex")
```

# Qualtrics setup

We will send the invites to 75,987 GCF applicants.

- 72,382 emails 
- 68,126 SMS texts 

replace NA emails with fake (madeup) emails 

```{r}
library(glue)

pseudo_emails <- glue("{1:3605}listexp@gmail.com")

sample$email_address[sample$contact_method == "phone"] <- pseudo_emails

sample_augmented <- sample

sample_augmented_test <- sample_augmented %>% 
  group_by(contact_method, group, assignment) %>% 
  sample_n(size = 1, replace = FALSE)

sample_augmented_rest <- sample_augmented %>%
    filter(!(calfresh_application_id %in% sample_augmented_test$calfresh_application_id))

write_csv(sample_augmented, "sample_augmented.csv")

sample_augmented_rest <- read.csv("sample_augmented_rest.csv");
sample_augmented_test <- read.csv("sample_augmented_test.csv")

sample_augmented_test <- sample_augmented_test %>%
    mutate(phone_number = ifelse(!is.na(phone_number), paste0("1", phone_number), phone_number)) %>%
    mutate(phone_number = ifelse(!is.na(phone_number), substr(phone_number, nchar(phone_number) - 10, nchar(phone_number)), phone_number))

sample_augmented_rest <- sample_augmented_rest %>%
    mutate(phone_number = ifelse(!is.na(phone_number), paste0("1", phone_number), phone_number)) %>%
    mutate(phone_number = ifelse(!is.na(phone_number), substr(phone_number, nchar(phone_number) - 10, nchar(phone_number)), phone_number))
```